Application of Micro-Rotors to Induce Crystal Nucleation: Challenges, Takeaways, and Future Prospects\\ PhD Thesis

### 2025-01-26 Latex_files/Thesis/chapters/chapter 2/main.tex
updates

\chapter{Theory and methods}
\label{chapter:theory}
The two key areas of this project cover two widely different
fields; crystallisation theory, which is covered in the 
previous introduction section, and tweezing theory. The 
following chapter summarises the working principles behind 
optical tweezing, the differences in scattering theories, 
and the broader theoretical calculations used throughout the 
rest of the thesis.

@@

Proper understanding of optical tweezing requires a description
of how the trapped particle interacts with the trapping laser. 
From an electromagnetism perspective the laser creates a 
spatially and temporally coherent electric field that scatters 
light off of a trapped particle. The laws governing electric 
and magnetic fields are summarised most succinctly via the 
Maxwell equations. The differential forms of which are given 
below \cite{Jackson_1975}:
\begin{align}
  \nabla \cdot \mathbf{E}
  &= \frac{\rho_ v}{\epsilon_0}
  \\
  \nabla \cdot \mathbf{B}
  &= 0
  \\
  \nabla \times \mathbf{E}
  &= \frac{\partial \mathbf{B}}{\partial t}
  \\
  \nabla \times \mathbf{B}
  &= {\bf j} +\frac{\partial \mathbf{E}}{\partial t}    
\end{align}
where $\rho_v$ is the free charge density, ${\bf j}_0$ is the 
charge current, and $\epsilon_0$ is the permittivity of free 
space.

@@

Where $\mathbf{P}$ is the polarization field and $\mathbf{M}$ 
is the magnetisation field, these fields arise due to the 
bound charges throughout the medium interacting with the EM 
field.

@@

The scattering force arises from the scattered field of the 
trapping beam pushing the target object away from the focal 
point. While some trapping arrangements can induce transverse 
scattering forces for simple spheres the scattering force is 
only significant in the direction of propagation \cite{Capitanio2002}.
The equilibrium position is found when the gradient force 
cancels out the the scattering force, and the thermal energy 
of the particle.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Harmonic Traps}
\label{sec:harmonic_traps}
When a sphere is located at the centre of a laser the gradient 
force will try to recentre the sphere if it is displaced. If 
the restoring force is proportional to the sphere's displacement 
this is referred to as a harmonic trap. The ratio of the restoring 
force to the displacement is called the trap stiffness (denoted 
by $\kappa$), and is an important characteristic of the trap. Fig.~\ref{fig:harmonic_trap} demonstrates an example case of the
force-displacement curve for a simple sphere in an optical trap. 
Laser power and numerical aperture are omitted as the exact shape 
is not of importance, only the linear region about the origin.
\begin{figure}[h!]
	\centering
	\includegraphics[width=1.1\linewidth]{sphere_force_curve.png}
	\caption{Example of a sphere's ($r=1\ \mu m$) force-displacement curve 
			while moving along the x-axis, where $x=0$ is the centre of the
			beam. The harmonic regime is highlight by the black linear fit.}
	\label{fig:harmonic_trap}
\end{figure}

@@

The Lorenz-Mie theory provides an exact solution to the Maxwell 
equations for the scattering caused by an isotropic sphere. This 
theory describes the scattered wave given off by a dielectric 
sphere when incident by a plane wave as a summation of partial 
spherical waves. For any spherical wave the vector fields must 
solve the Helmholtz wave equation given by:
\begin{align}
	\nabla^2\mathbf{E} +k^2\mathbf{E} = 0
\end{align}

Where $k$ is the wave number of the electromagnetic radiation 
($k = 2\pi/\lambda$). This combined with the constraints of 
Maxwell's equations leaves very few exact solutions apart from 
spherical or planar waves. For a laser of arbitrary shape the 
electric field can be described as a summation of partial 
spherical waves multiplied by an expansion coefficient. Depending 
on the beam in question the expansion coefficients can be tailored
to match the desired beam shape. For a approximate Gaussian beam 
interacting with a particle of volume $V$, the incident, internal, 
and scattered fields are given as \cite{RanhaNeves2019}: 
\begin{align}
	\label{eq:incident}
	{\bf E}_{\rm inc}({\bf r})
	&=
	E_0 \sum_{n}^\infty\sum_{m=-n}^n \left[
	a_{nm}{\bf M}_{nm}^{(1)}({\bf r})
	+ b_{nm}{\bf N}_{nm}^{(1)}({\bf r})\right]
	\\
	\label{eq:internal}
	{\bf E}_{\rm int}({\bf r})
	&=
	E_0 \sum_{n}^\infty \sum_{m=-n}^{n}\left[
	d_{nm}{\bf N}_{nm}^{(1)}({\bf r}) + c_{nm}{\bf M}_{nm}^{(1)}({\bf r})
	\right]
	\\
	\label{eq:scattered}
	{\bf E}_{\rm scat}({\bf r})
	&=
	E_0 \sum_{n}^\infty \sum_{m=-n}^n\left[
	p_{nm}{\bf N}_{nm}^{(3)}({\bf r})+q_{nm}{\bf M}_{nm}^{(3)}({\bf r})
	\right] 
\end{align}

where $a_{nm}$, $b_{nm}$, $c_{nm}$, $d_{nm}$, $p_{nm}$, and $q_{nm}$
are the expansion coefficients of each of the fields, and ${\bf M}_
{nm}$ and ${\bf N}_{nm}$ are the magnetic and electric vector 
spherical harmonics respectively. The superscripts denote whether 
the field is inward [(1)] or outward propagating [(3)]. By linearity 
of Maxwell's equations, its possible to relate the expansion 
coefficients of the scattered field directly to those of the incident 
field. As such we can write the scattering coefficients in terms of 
their incident components and the target's 'T-matrix'.
\begin{align}
	\label{eq:t_matrix}
	\begin{pmatrix}
		q_{mn} \\
		p_{mn} 
	\end{pmatrix}
	= \bold{T} 
	\begin{pmatrix}
		a_{mn} \\
		b_{mn}
	\end{pmatrix}
	= \begin{pmatrix}
		T_{11} & T_{12} \\
		T_{21} & T_{22}
	\end{pmatrix}
	\begin{pmatrix}
		a_{mn} \\
		b_{mn}
	\end{pmatrix}
\end{align}

@@

In order to compute the optical force and torque applied to a 
given object one simply needs to know the momentum carried by 
the incident and scattered fields. So long as we know the 
expansion coefficients of the incident beam we can find the 
scattered beam coefficients via Eq.~\eqref{eq:t_matrix} to 
compute the force. This project utilises the results from 
Farsund \cite{Farsund1996} but in the form used by Crichton 
\cite{Crichton2000THEMD}. As an example the time averaged linear 
momentum carried by any beam can be computed via:
\begin{equation}
	\label{eq:linear_momentum_in}
	\begin{split}
		\langle P_{z,\ in} \rangle
		&=
		-\frac{1}{4\pi k^2}\sum_{n,m} \left(\frac{m}{n(n+1)}\Im[a_{nm}b^*_{nm}] \right.
		\\ 
		&+\frac{1}{n+1}\left[\frac{n(n+2)(n-m+1)(n+m+1)}{(2n+1)(2n+3)} \right]^{1/2}
		\\
		& \left.\times\Im[b_{nm}b^*_{nm}+a_{nm}a^*_{n+1m}] \right)
	\end{split}
\end{equation}

@@

\noindent
Where $a_{nm}, b{nm}, p_{nm}, \& q_{nm}$ are the beam expansion 
coefficients used in \eqref{eq:incident} \& \eqref{eq:scattered}.
Similar expressions are shown in Sec.~\ref{sec:simulations} for 
calculating the optical torque. Lorenz-Mie theory can be applied 
to describe the scattering from any particle regardless of size, 
though as the size of the target particle changes - relative to 
the wavelength of light - the theory describing the electromagnetic 
fields are easier to approximate by alternative theories. When the 
particle radius is $\gg \lambda$ the optical force is best described 
by the Ray-Optics model. In the opposite case the target particle is 
better approximated as a single electric dipole as described by 
the Rayleigh approximation.

@@

\noindent 
where $Q_i$ is the trapping efficiency, $\Delta n$ is the difference
in refractive indices between the solution and the target particle,
and $P_i$ is the power of the individual ray. The net force can be 
subdivided into its gradient and scattering components, where the latter
directs the particle to the centre of the beam, and the former acts on 
the particle in the direction of beam propagation. For a beam with a
Gaussian intensity distribution $P_i$ will fall off as you move from
the centre of the beam. The ray optics model is ideal when dealing 
with larger particles whose diameter far exceed the wavelength of light
being scattered. While it can be useful in predicting the forces 
experienced by said particle's it does not fully capture the behaviour
of light when considering interference between different rays.

@@

The Rayleigh approximation is for describing a particle when 
the particle radius is $\ll \lambda$. The underlying theory 
is that a dielectric sphere can be treated as a dipole while 
in the presence of the electromagnetic field. In which case 
the scattering force is given simply by the scattering of 
the induced dipole, and the gradient force is due to the 
Lorentz force \cite{Gordon1973}. The gradient forces in the 
principle Cartesian axis are described by Harada et al \cite{YasuhiroHarada1996} in MKS units as a restorative 
rectangular force field:
\begin{align}
  F_{grad,x}
  &=-\hat{x} \frac{2\pi n_2 a^3}{c}
    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{x}/w_0}{1+(2\tilde{z})^2} \times I(r)
  \label{eq:Rayleigh_start}\\
  F_{grad,y}
  &=-\hat{y} \frac{2\pi n_2 a^3}{c}
    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{y}/w_0}{1+(2\tilde{z})^2} \times I(r)
  \\
  F_{grad,z}
  &=-\hat{z} \frac{2\pi n_2 a^3}{c}
    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{z}/w_0}{1+(2\tilde{z})^2}
    \nonumber \times I(r)
  \\ 
  & \times \left[1-\frac{2(\tilde{x}^2+\tilde{y}^2)}{1+(2\tilde{z})^2} \right]
  \\
  \label{eq:Rayleigh_end}
  \text{where:}
  \nonumber
  \\
	I(r) &= \left(\frac{2P}{\pi w_0^2}\right) \frac{1}{1+(2\tilde{z}^2)} 
	\exp \left[ - \frac{2(\tilde{x}^2+\tilde{y}^2)}{1+(2\tilde{z})^2} \right]
\end{align}

@@

There is a clear difference between Eq.~\eqref{eq:Rayleigh_start}-
\eqref{eq:Rayleigh_end} and Eq.~\eqref{eq:Fz}, namely that the 
Rayleigh regime is not concerned with the state of the scattered
field. This is because if the particle is approximated by a point
dipole we can ignore the momentum transfer calculation and simply
compute the Lorenz force for a dipole in an electric field. While 
not important for optical force calculation higher complexity 
scattering problems can be simplified by subdividing the particle 
into discrete dipoles (see Sec~\ref{sec:scattering}), in which case 
the scattered field is now significant. As the target particle gets 
larger this approximation fails to accurately describe the trapping force 
\cite{Li2021}. For larger particles the optical force and scattering 
is better described via the Lorenz-Mie theory.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Scattering methods}
\label{sec:scattering}
There are several methods available to calculate the scattered 
field produced by a particle. While one can compute this by 
directly applying Lorenz-Mie theory, describing the interactions 
between focused beams and complicated shapes can be computationally 
slow and memory intensive. Several alternative scattering approaches 
have been developed that allow for faster computation times 
simplifying the scattering problem. While all of these methods will
produce similar results the key difference is in how one represents
the scatterer.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{T-matrix Method}
The $T$-matrix method was first developed by Peter Waterman with 
his research into acoustic wave scattering \cite{Waterman1969}, 
this would later be extended to electromagnetic waves. Sometimes 
referred to as the extended boundary condition method (ECBM), the 
method replaces the scatterer with a series of surface currents 
over the targets surface. These currents are chosen so that the 
electromagnetic field outside is identical to the original problem 
\cite{Wriedt1998}. By expanding the harmonic functions one can 
derive the individual elements of \textbf{T} until a desired 
accuracy is achieved. One can either use the ECBM to compute the 
scattering for a fixed particle orientation and position. Or they
can evaluate the scattering for any particle orientation by 
computing its $T$-matrix.

For this project we utilised an extension of the ECBM known as 
multi-sphere $T$-matrix method (\textit{mstm}) developed by 
Mackowski \cite{Mackowski2011}. Where Lorenz-Mie Theory centres
the spherical harmonics expansion about the sphere's origin it
often struggles to adjust the expansion to any arbitrary origin.
\textit{mstm} utilises an addition theorem to translate the 
individual spherical expansions from the centres of each 
individual sphere to a preset target origin. \textit{mstm}'s 
primary function is to compute the averaged scattering from 
large spherical aggregates, but for our purposes it can be
also utilised to compute $T$ matrices for arbitrary shaped
spherical aggregates.

The $T$-matrix method is exceptionally useful for computing the 
scattering from any arbitrary spherical aggregate. However, the 
$T$-matrix method by itself can be computationally taxing as the 
number of spheres increases \cite{Mackowski2011}. While it is 
possible to solve for the full electromagnetic field of each the 
entire cluster this is only applicable for a single orientation 
and can be even slower for large aggregates \cite{Mackowski1996, 
Xu1995}. The benefit of \textit{mstm} is that the major scattering 
properties (scattering and extinction cross sections, scattering 
matrices, etc) can all be computed both for single orientations, 
or averaged over multiple orientations to determine the average 
scattering from the target particle.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Discrete Dipole Approximation}
The discrete dipole approximation (DDA) is a general method that 
can be applied to the scattering from particles of arbitrary 
composition and geometry. Developed by Purcell and Pennypacker 
\cite{Purcell1973}, the DDA method approximates the particle as 
being constructed of discrete dipoles. This is essentially an 
extension of the Rayleigh approximation, though in order to compute 
the optical force one requires a full description of the electric 
field. Each dipole interacts with both the incident field and the 
scattered fields from every other dipole surrounding it. The 
resulting scattered field is identical to the scattered field 
produced by direct integration of Eq.~\eqref{eq:internal} 
throughout the full particle volume \cite{Goedecke1988}. The 
integral form for the total electric field inside a scatterer is 
given as \cite{Wriedt1998}:
\begin{align}
	\mathbf{E(r)} = \mathbf{E}_{inc}(r) + \sum^{N_d}_{i}\mathbf{E}_{scat,i} =  
	\mathbf{E}_{inc}(r) + \int_{V/V_0}d^3r\mathbf{\bar{G}(r,r')}
	\chi(\mathbf{r'})\mathbf{E(r')}
	\label{eq:DDA}
\end{align}

Where $\mathbf{\bar{G}}$ is the Greens dyadic function of free space, 
which defines the impulse response between two separate dipoles; and 
$\chi$ is the susceptibility of the medium, which describes the 
degree of polarisation of the medium in the presence of an electric 
field. One of the primary advantages of DDA over the T-matrix method 
is that the composition of the target can be changed freely. When 
comparing different computational scattering methods, the ECBM was 
found to be better suited for simulating the scattering of symmetric
targets as the ECBM can use the target's symmetry to speed up 
calculations \cite{Wriedt1998}. But when dealing with inhomogeneous 
media DDA is more efficient compared to ECBM.

The DDA method can be used to derive the target's $T$-matrix. After
generating the far field expansion of the electric field (Eq.~
\eqref{eq:DDA}), the vector spherical expansion far field 
(Eq.~\eqref{eq:scattered}) can be matched to it. Solving for the 
expansion coefficients $p_{nm}$, and $q_{nm}$ allows one to construct
the T-matrix of the target. 
\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Describing the trajectory of a trapped particle}
While describing the optical forces is sufficient for 
theoretical characterisation of an optical trap, trying 
to replicate that model in an experimental setting is 
near impossible. As such it is often far simpler to 
characterise the optical trap by recording the particle's
trajectory (see \ref{sec:position_detection} for 
specifics regarding the equipment), and then fitting it 
to our presumed description of the beam. The trajectory 
is easily understood if we understand the particle's 
diffusive behaviour.

\section{Langevin Equation}
Describing any microscopic motion requires an understanding 
of a molecules diffusive behaviour, for the case of optical 
tweezers the most complete model of diffusion is the Langevin 
equation. Models such as the Fickian, and Einstein derivations 
are sufficient for macroscopic behaviours the Langevin 
equation better describes the microscopic characteristics of 
any diffusive behaviour.

>>>>

\chapter{Theory and methods}
\label{chapter:theory}
The two key areas of this project cover two somewhat unconnected
fields; crystallisation theory, which is covered in the 
previous introduction section, and tweezing theory. The 
following chapter summarises the working principles behind 
optical tweezing, the differences in scattering theories, 
and the broader theoretical calculations used throughout the 
rest of the thesis.

@@

Proper understanding of optical tweezing requires a description
of how the trapped particle interacts with the laser. 
From an electromagnetism perspective the laser creates a 
spatially and temporally coherent electric field that scatters 
light when passing though a medium. The laws governing electric 
and magnetic fields are summarised most succinctly via the 
Maxwell equations. The differential forms of which are given 
below \cite{Jackson_1975}:
\begin{align}
  \nabla \cdot \mathbf{E}
  &= \frac{\rho_ v}{\epsilon_0}
  \label{eq:max_start}
  \\
  \nabla \cdot \mathbf{B}
  &= 0
  \\
  \nabla \times \mathbf{E}
  &= \frac{\partial \mathbf{B}}{\partial t}
  \\
  \nabla \times \mathbf{B}
  &= {\bf j} +\frac{\partial \mathbf{E}}{\partial t}
  \label{eq:max_end}    
\end{align}
where $\rho_v$ is the free charge density, ${\bf j}_0$ is the 
charge current, and $\epsilon_0$ is the permittivity of free 
space.

@@

Where $\mu_0$ is the vacuum permeability $\mathbf{P}$ 
is the polarization field and $\mathbf{M}$ is the 
magnetisation field, these fields arise due to the 
bound charges throughout the medium interacting with 
the EM field.

@@

The scattering force arises from the motion of bound charges 
generating secondary scattering as they move through the 
electric field \cite{YasuhiroHarada1996}. Even if the particle
experiences no external forces the scattering force is still 
significant as the electric field oscillates, though for most
practical applications the time dependence is ignored. While 
some trapping arrangements can induce transverse scattering 
forces for simple spheres the scattering force is only 
significant in the direction of propagation \cite{Capitanio2002}.
In order to determine where a particle will remain at rest 
in an optical trap requires determining where the potential 
well formed by the gradient force is equal to the thermal 
energy of the system.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Harmonic Traps}
\label{sec:harmonic_traps}
When a sphere is located at the centre of a laser the gradient 
force will try to recentre the sphere if it is displaced. If 
the restoring force is proportional to the sphere's displacement 
this is referred to as a harmonic trap. The ratio of the restoring 
force to the displacement is called the trap stiffness (denoted 
by $\kappa$), and is an important characteristic of the trap. Fig.~\ref{fig:harmonic_trap} demonstrates an example case of the
force-displacement curve for a simple sphere in an optical trap. 
Laser power and numerical aperture are omitted as the exact shape 
is not of importance, only the linear region about the origin.
\begin{figure}[h!]
	\centering
	\includegraphics[width=1.1\linewidth]{sphere_force_curve.png}
	\caption{Example of a sphere's ($r=1\ \mu m$) force-displacement
		curve while moving along the x-axis, where $x=0$ is the 
		centre of the beam. The harmonic regime is highlight by the 
		linear fit given in black.}
	\label{fig:harmonic_trap}
\end{figure}

@@

The Lorenz-Mie theory provides an exact solution to the Maxwell 
equations for the scattering caused by an isotropic sphere. This 
theory describes the scattered wave given off by a dielectric 
sphere when incident by a plane wave as a summation of partial 
spherical waves. For any wave equation the vector fields must 
solve the Helmholtz equation given by:
\begin{align}
	\nabla^2\mathbf{E} +k^2\mathbf{E} = 0
\end{align}

Where $k$ is the wave number of the electromagnetic radiation 
($k = 2\pi/\lambda$). This combined with the constraints of 
Maxwell's equations leaves very few exact solutions apart from 
spherical or planar waves. For a laser of arbitrary shape the 
electric field can be described as a summation of partial 
spherical waves multiplied by an expansion coefficient. Depending 
on the beam in question the expansion coefficients can be tailored
to match the desired beam shape. For a approximate Gaussian beam 
interacting with a sphere of radius $a$, the incident, internal, 
and scattered fields are given as \cite{RanhaNeves2019}: 
\begin{align}
	\label{eq:incident}
	{\bf E}_{\rm inc}({\bf r})
	&=
	E_0 \sum_{n}^\infty\sum_{m=-n}^n \left[
	a_{nm}{\bf M}_{nm}^{(1)}({\bf r})
	+ b_{nm}{\bf N}_{nm}^{(1)}({\bf r})\right]
	\\
	\label{eq:internal}
	{\bf E}_{\rm int}({\bf r})
	&=
	E_0 \sum_{n}^\infty \sum_{m=-n}^{n}\left[
	d_{nm}{\bf N}_{nm}^{(1)}({\bf r}) + c_{nm}{\bf M}_{nm}^{(1)}({\bf r})
	\right]
	\\
	\label{eq:scattered}
	{\bf E}_{\rm scat}({\bf r})
	&=
	E_0 \sum_{n}^\infty \sum_{m=-n}^n\left[
	p_{nm}{\bf N}_{nm}^{(3)}({\bf r})+q_{nm}{\bf M}_{nm}^{(3)}({\bf r})
	\right] 
\end{align}

where $a_{nm}$, $b_{nm}$, $c_{nm}$, $d_{nm}$, $p_{nm}$, and $q_{nm}$
are the expansion coefficients of each of the fields, and ${\bf M}_
{nm}$ and ${\bf N}_{nm}$ are the magnetic and electric vector 
spherical harmonics respectively. The superscript denote whether 
the field is inward [(1)] or outward propagating [(3)]. By linearity 
of Maxwell's equations, its possible to relate the expansion 
coefficients of the scattered field directly to those of the incident 
field. As such we can write the scattering coefficients in terms of 
their incident components and the target's 'T-matrix'.
\begin{align}
	\label{eq:t_matrix}
	\begin{pmatrix}
		q_{mn} \\
		p_{mn} 
	\end{pmatrix}
	= \bold{T} 
	\begin{pmatrix}
		a_{mn} \\
		b_{mn}
	\end{pmatrix}
	= \begin{pmatrix}
		T_{11} & T_{12} \\
		T_{21} & T_{22}
	\end{pmatrix}
	\begin{pmatrix}
		a_{mn} \\
		b_{mn}
	\end{pmatrix}
\end{align}

@@

In order to compute the optical force and torque applied to a 
given object one simply needs to know the momentum carried by 
the incident and scattered fields. So long as we know the 
expansion coefficients of the incident beam we can find the 
scattered beam coefficients via Eq.~\eqref{eq:t_matrix} to 
compute the force. This project utilises the results from 
Farsund \cite{Farsund1996} but in the form used by Crichton 
\cite{Crichton2000THEMD}. As an example the time averaged linear 
momentum carried by any incident beam can be computed via:
\begin{equation}
	\label{eq:linear_momentum_in}
	\begin{split}
		\langle P_{z,\ in} \rangle
		&=
		-\frac{1}{4\pi k^2}\sum_{n,m} \left(\frac{m}{n(n+1)}\Im[a_{nm}b^*_{nm}] \right.
		\\ 
		&+\frac{1}{n+1}\left[\frac{n(n+2)(n-m+1)(n+m+1)}{(2n+1)(2n+3)} \right]^{1/2}
		\\
		& \left.\times\Im[b_{nm}b^*_{nm}+a_{nm}a^*_{n+1m}] \right)
	\end{split}
\end{equation}

@@

\noindent
Where $a_{nm}, b_{nm}, p_{nm}, \& q_{nm}$ are the beam expansion 
coefficients used in \eqref{eq:incident} \& \eqref{eq:scattered}.
Similar expressions are shown in Sec.~\ref{sec:simulations} for 
calculating the optical torque. Lorenz-Mie theory can be applied 
to describe the scattering from any particle regardless of size.
Though as the size of the target particle changes - relative to 
the wavelength of light - the theory describing the electromagnetic 
fields are easier to approximate by alternative theories. When the 
particle radius is $\gg \lambda$ the optical force is best described 
by the Ray-Optics model. In the opposite case the target particle is 
better approximated as a single electric dipole as described by 
the Rayleigh approximation.

@@

\noindent 
where $Q_i$ is the trapping efficiency, $\Delta n$ is the difference
in refractive indices between the solution and the target particle,
and $P_i$ is the power of the individual ray. The net force can be 
subdivided into its gradient and scattering components, where the latter
directs the particle to the centre of the beam, and the former acts on 
the particle in the direction of beam propagation. For a beam with a
Gaussian intensity distribution $P_i$ will fall off as you move from
the centre of the beam. The ray optics model is ideal when dealing 
with larger particles whose diameter far exceed the wavelength of light
being scattered. While it can be useful in predicting the forces 
experienced by said particle's it does not fully capture the behaviour
of light when considering interference between different rays. And 
as such no information is learned about the scattered field.

@@

The Rayleigh approximation is for describing a particle when 
the particle radius is $\ll \lambda$. The underlying theory 
is that a dielectric sphere can be treated as a dipole while 
in the presence of the electromagnetic field. In which case 
the scattering force is given simply by the scattering of 
the induced dipole, and the gradient force is due to the 
Lorentz force \cite{Gordon1973}. The gradient forces in the 
principle Cartesian axis are described by Harada et al \cite{YasuhiroHarada1996} in MKS units as a restorative 
rectangular force field:
\begin{align}
  F_{grad,x}
  &=-\hat{x} \frac{2\pi n_2 a^3}{c}
    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{x}/w_0}{1+(2\tilde{z})^2} \times I(r)
  \label{eq:Rayleigh_start}\\
  F_{grad,y}
  &=-\hat{y} \frac{2\pi n_2 a^3}{c}
    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{y}/w_0}{1+(2\tilde{z})^2} \times I(r)
  \\
  F_{grad,z}
  &=-\hat{z} \frac{2\pi n_2 a^3}{c}
    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{z}/w_0}{1+(2\tilde{z})^2}
    \nonumber \times I(r)
  \\ 
  & \times \left[1-\frac{2(\tilde{x}^2+\tilde{y}^2)}{1+(2\tilde{z})^2} \right]
  \\
  \label{eq:Rayleigh_end}
\end{align}
where, for a approximate Gaussian beam the intensity $I(r)$ 
close to the beam focus can be calculated as such.
\begin{align}
  \nonumber
	I(r) &= \left(\frac{2P}{\pi w_0^2}\right) \frac{1}{1+(2\tilde{z}^2)} 
	\exp \left[ - \frac{2(\tilde{x}^2+\tilde{y}^2)}{1+(2\tilde{z})^2} \right]
\end{align}

@@

There is a clear difference between Eq.~\eqref{eq:Rayleigh_start}-
\eqref{eq:Rayleigh_end} and Eq.~\eqref{eq:linear_momentum_in}- \eqref{eq:Fz}, namely that the Rayleigh regime is not concerned with 
the state of the scattered field. This is because if the particle 
is approximated by a point dipole we can ignore the momentum 
transfer calculation and simply compute the Lorenz force for a 
dipole in an electric field. While not important for optical force calculation higher complexity scattering problems can be simplified 
by subdividing the particle into discrete dipoles (see Sec~\ref{sec:scattering}), in which case the scattered field is now significant. As the target particle gets larger this approximation 
fails to accurately describe the trapping force \cite{Li2021}. For 
larger particles the optical force and scattering is better described 
via the Lorenz-Mie theory.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Scattering methods}
\label{sec:scattering}
There are several methods available to calculate the scattered 
field produced by a particle. While one can compute this by 
directly applying Lorenz-Mie theory, describing the interactions 
between focused beams and complicated shapes can be computationally 
slow and memory intensive. Several alternative scattering approaches 
have been developed that allow for faster computation times 
simplifying the scattering problem. While all of these methods will
produce similar results the key difference is in how one represents
the scatterer.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{T-matrix Method}
The $T$-matrix method was first developed by Peter Waterman with 
his research into acoustic wave scattering \cite{Waterman1969}, 
this would later be extended to electromagnetic waves. Sometimes 
referred to as the extended boundary condition method (EBCM), the 
method replaces the scatterer with a series of surface currents 
over the targets surface. These currents are chosen so that the 
electromagnetic field outside is identical to the original problem 
\cite{Wriedt1998}. By expanding the harmonic functions one can 
derive the individual elements of \textbf{T} until a desired 
accuracy is achieved. One can either use the EBCM to compute the 
scattering for a fixed particle orientation and position. Or they
can evaluate the scattering for any particle orientation by 
computing its $T$-matrix. Because the T-matrix is based only 
on the properties of the scatterer it can be reused in 
calculations regardless of the incident field in question.

For this project we utilised an extension of the EBCM known as 
multi-sphere $T$-matrix method (\textit{mstm}) developed by 
Mackowski \cite{Mackowski2011}. Where Lorenz-Mie Theory centres
the spherical harmonics expansion about the sphere's origin it
often struggles to adjust the expansion to any arbitrary origin.
\textit{mstm} utilises an addition theorem to translate the 
individual spherical expansions from the centres of each 
individual sphere to a preset target origin. \textit{mstm}'s 
primary function is to compute the averaged scattering from 
large spherical aggregates, but for our purposes it can be
also utilised to compute $T$ matrices for arbitrary shaped
spherical aggregates.

The $T$-matrix method is exceptionally useful for computing the 
scattering from any arbitrary spherical aggregate. However, the 
$T$-matrix method by itself can be computationally taxing as the 
number of spheres increases \cite{Mackowski2011}. While it is 
possible to solve for the full electromagnetic field for the 
entire cluster this is only applicable for a single orientation 
and can be even slower for large aggregates \cite{Mackowski1996, 
Xu1995}. The benefit of \textit{mstm} is that the major scattering 
properties (scattering and extinction cross sections, scattering 
matrices, etc) can all be computed both for single orientations, 
or averaged over multiple orientations to determine the average 
scattering from the target particle.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Discrete Dipole Approximation}
The discrete dipole approximation (DDA) is a general method that 
can be applied to the scattering from particles of arbitrary 
composition and geometry. Developed by Purcell and Pennypacker 
\cite{Purcell1973}, the DDA method approximates the particle as 
being constructed of discrete dipoles. This is essentially an 
extension of the Rayleigh approximation, though in order to compute 
the optical force one now requires a full description of the electric 
field. Each dipole interacts with both the incident field and the 
scattered fields from every other dipole surrounding it. The 
resulting scattered field is identical to the scattered field 
produced by direct integration of Eq.~\eqref{eq:internal} 
throughout the full particle volume \cite{Goedecke1988}. The 
integral form for the total electric field inside a scatterer is 
given as \cite{Wriedt1998}:
\begin{align}
	\mathbf{E(r)} = \mathbf{E}_{inc}(r) + \sum^{N_d}_{i}\mathbf{E}_{scat,i} =  
	\mathbf{E}_{inc}(r) + \int_{V/V_0}d^3r\mathbf{\bar{G}(r,r')}
	\chi(\mathbf{r'})\mathbf{E(r')}
	\label{eq:DDA}
\end{align}

Where $\mathbf{\bar{G}}$ is the Greens dyadic function of free space, 
which defines the impulse response between two separate dipoles; and 
$\chi$ is the susceptibility of the medium, which describes the 
degree of polarisation of the medium in the presence of an electric 
field. One of the primary advantages of DDA over the T-matrix method 
is that the composition of the target can be changed freely. When 
comparing different computational scattering methods, the EBCM was 
found to be better suited for simulating the scattering of symmetric
targets as the EBCM can use the target's symmetry to speed up 
calculations \cite{Wriedt1998}. But when dealing with inhomogeneous 
media DDA is more efficient compared to EBCM.

The DDA method can be used to derive the target's $T$-matrix. After
generating the far field expansion of the electric field (Eq.~
\eqref{eq:DDA}), the vector spherical expansion far field 
(Eq.~\eqref{eq:scattered}) can be matched to it. Solving for the 
expansion coefficients $p_{nm}$, and $q_{nm}$ allows one to construct
the T-matrix of the target. 
\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Describing the trajectory of a trapped particle}
While describing the optical forces is sufficient for 
theoretical characterisation of an optical trap, in an 
experimental setting the observed motion is not as 
simple. As such it is often far simpler to 
characterise the optical trap by recording the particle's
trajectory (see \ref{sec:position_detection} for 
specifics regarding the equipment), and then fitting it 
to our presumed description of the beam. The trajectory 
is easily understood if we understand the particle's 
diffusive behaviour.

\subsection{Langevin Equation}
Describing any microscopic motion requires an understanding 
of a molecules diffusive behaviour, for the case of optical 
tweezers the most complete model of diffusion is the Langevin 
equation. Models such as the Fickian, and Einstein derivations 
are sufficient for macroscopic behaviours but the Langevin 
model better describes the microscopic characteristics of 
any diffusive behaviour.
